Making Sequential Programs Deterministic

### 2019-03-28 design.tex
iteration rules
todo: type rules

 Although such types could exist, they are
disallowed to prevent the following bug:

@@

A collection or map annotated as \<OrderNonDet> has the following properties:
\todo{Write these as type rules.}
\begin{enumerate}
    \item The individual elements extracted by iterating over the collection get the type \<NonDet>.
    \item Accessing properties such as \<size()> or querying if the collection \<isEmpty()> will return types
    that are annotated as \<Det> as they do not depend on the iteration order. 
\end{enumerate}

>>>>

 Although \<NonDet> collections containing \<Det> or \<OrderNonDet> elements could exist, they are
disallowed to prevent the following bug:

@@

A collection or map annotated as \<OrderNonDet> has the following properties:
%\todo{Write these as type rules.}
\begin{enumerate}
    \item The individual elements extracted by iterating over the collection get the type \<NonDet> (Rules \rulename{iteration} and \rulename{next element} in Figure~\ref{fig-ordernondet-rules}). 
    Rule \rulename{hasnext element} states that querying if the iterator of an \<OrderNonDet Collection> \<hasNext()> will result in a \<Det boolean>.
    \item Accessing properties such as \<size()> or querying if the collection \<isEmpty()> will return types
    that are annotated as \<Det> as they do not depend on the iteration order. 
\end{enumerate}

\begin{figure}
    $\infer[\rulename{iteration}]
    {\Gamma \vdash \|y| : \|OrderNonDet\ Iterator|\angles{\tau}}
    {\Gamma \vdash \|x| : \|OrderNonDet \ C\angles{\tau}|
    & C \sqsubseteq \|Collection|
    & \|y = C.iterator()|
    & \wellformed{\tau}
    }$

    $\infer[\rulename{next element}]
    {\Gamma \vdash \|z| : \|NonDet \ \beta|}
    {\Gamma \vdash \|y| : \|OrderNonDet \ Iterator\angles{\kappa \ \beta}|
        & \|z = y.next()|
        & \wellformed{\kappa \ \beta}
    }$

    $\infer[\rulename{hasnext element}]
    {\Gamma \vdash \|z| : \|Det \ boolean|}
    {\Gamma \vdash \|y| : \|OrderNonDet \ Iterator\angles{\tau}|
        & \|z = y.hasNext()|
        & \wellformed{\tau}
    }$

    $\infer[\rulename{size}]
    {\Gamma \vdash \|y| : \|Det \ int|}
    {\Gamma \vdash \|x| : \|OrderNonDet \ C\angles{\tau}|
        & C \sqsubseteq \|Collection|
        & \|y = x.size()|
        & \wellformed{\tau}
    }$
\caption{\<OrderNonDet> Collection rules}
\label{fig-ordernondet-rules}
\end{figure}
